Practical Machine Learning Project
========================================================

Background
========================================================
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

Data
========================================================
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

Goal and analysis
========================================================
The goal of this project is determine or predict the manner in wich the partcipants did exercise.

So, first I load the packages for start the study:
```{r, message=FALSE, warning=FALSE}
library(AppliedPredictiveModeling)
library(caret)
library(ElemStatLearn)
library(pgmm)
library(rpart)
library(randomForest)
library(rattle)
library(ggplot2)
library(e1071)
library(rpart.plot)
```
Obtain the data set, train and test, and fixed a seed.
```{r, message=FALSE, warning=FALSE}
set.seed(3633)

urlTrain<-"http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
urlTest<-"http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

train_data<-read.csv(urlTrain, header=T, na.strings=c("NA","#DIV/0!",""))
test_data<-read.csv(urlTest, header=T, na.strings=c("NA","#DIV/0!",""))

```
We make a small exploratory data analysis
```{r, echo=TRUE, message=FALSE, warning=FALSE}
dim(train_data)
attributes(train_data)$names
attributes(test_data)$names==attributes(train_data)$names
```
Clean data for this column having 90% of NAs will be removed. In the same way the variables X, user, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window and num_window are deleted because they do not provide meaningful information to the analysis.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
PSum_var<-(colSums(is.na(train_data))/dim(train_data)[1])
PSum_var
var_nelim<-NULL
j=1
for(i in 8:dim(train_data)[2])
{
   if(PSum_var[i]<=.90)
   {
     var_nelim[j]=i
     j=j+1
   } 
}
```
varnelim represents the variables to be considered in the analysis, these variables are shown below
```{r}
var_nelim
```
Select those variables in train_data and test_data
```{r}
train_data<-train_data[,var_nelim]
test_data<-test_data[,var_nelim]
dim(test_data)
dim(train_data)
```
Divide the data, we selected 80% for the training data and 20% for test data
```{r, message=FALSE, warning=FALSE}
attach(train_data)
inTrain<-createDataPartition(classe, p=.8, list=F)
training<-train_data[inTrain,]
testing<-train_data[-inTrain,]
dim(training)
dim(testing)
```
We adjust a first model, a tree
```{r,message=FALSE, warning=FALSE}
fit_data<-rpart(classe~., data=training, method="class" )
pred1<-predict(fit_data, testing, type="class")
confusionMatrix(testing$classe,pred1)
```
The accuracy rate is 0.7545. 
```{r fig.width=7, fig.height=6, message=FALSE, warning=FALSE}
fancyRpartPlot(fit_data)
```
The second model a random forest
```{r,message=FALSE, warning=FALSE}
fitRF<-randomForest(classe~., data=training)
predRF<-predict(fitRF, testing, type="class")
confusionMatrix(testing$classe,predRF)
```
The accuracy rate is 0.9977
```{r fig.width=7, fig.height=6, message=FALSE, warning=FALSE}
plot(fitRF)
```
```{r fig.width=7, fig.height=6, message=FALSE, warning=FALSE}
varImpPlot(fitRF)
```

Prediction on testing data
========================================================
The accuracy rate of the random forest is greater than the tree, therefore the prediction in the test set will be done with the random forest.

```{r}
predict(fitRF, newdata=test_data)
```